Non-kissing and non-crossing complexes \\ for gentle algebras

### 2018-05-30 draft/accordionComplexSurfaces.tex
minor changes

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, xifthen, ae, aecompl, blkarray, pifont, mathtools, a4wide}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}

@@

\begin{definition}
A \defn{marked surface}~$\bar\surface \eqdef (\surface, M)$ is an orientable surface~$\surface$ with boundaries, together with a set~$M$ of marked points which can be on the boundary of~$\surface$ or not.
An \defn{arc} on~$\bar\surface$ is a curve on~$\surface$ connecting two marked points of~$M$ and whose interior is disjoint from~$M$ and the boundary of~$\surface$ (note that arcs homotopic to boundary arcs are allowed here).
As usual, arcs are considered up to homotopy relative to their endpoints in~$\surface \ssm M$.
Two arcs \defn{cross} when they intersect in their interior.
\end{definition}

\begin{definition}
A \defn{dissection} of~$\bar\surface$ is a collection~$\dissection$ of pairwise non-crossing arcs on~$\bar\surface$.
The \defn{faces} of a dissection~$\dissection$ are the connected components of the complement of the union of the arcs of~$\dissection$ in the surface~$\surface$.
We denote by~$\faces(\dissection)$ the set of faces of~$\dissection$.
The dissection~$\dissection$ is \defn{cellular} if all its faces are disks.
\end{definition}

@@

\begin{definition}
Let~$\surface$ be a surface with two sets~$M$ and~$M\dual$ of marked points so that~$M$ and~$M\dual$ alternate along the boundary of~$\surface$.
A cellular dissection~$\dissection$ of~$(\surface, M)$ and a cellular dissection~$\dissection\dual$ of~$(\surface, M\dual)$ are \defn{dual cellular dissections} if there are pairs of mutually inverse bijections~$M\dual \leftrightarrow \faces(\dissection)$ and~$M \leftrightarrow \faces(\dissection\dual)$, denoted~$\dual$ in both directions, such that~$\dissection$ has an arc joining its vertices~$u,v \in M$ and separating its faces~$f,g \in  \faces(\dissection)$ if and only if~$\dissection\dual$ has an arc joining its vertices~$f\dual, g\dual \in M\dual$ and separating its faces~$u\dual, v\dual \in \faces(\dissection\dual)$.
\end{definition}

Note that contrarily to the usual conventions, the dual vertex of a face~$f$ is not always in the interior of the face.
More precisely, there are two situations:
\begin{itemize}
\item if a face~$f$ has no arc on the boundary of~$\surface$, its dual vertex~$f\dual$ lies in the interior~of~$f$,
\item if a face~$f$ has an arc on the boundary of~$\surface$, then it has exactly one such arc and its dual vertex~$f\dual$ lies on this arc.
\end{itemize}
In fact, the second point forces the following characterization of the cellular dissections that admit a dual cellular dissection.

\begin{proposition}
\label{prop:conditionsDualDissections}
For a cellular dissection~$\dissection$ of a marked surface~$(\surface, M)$, the following assertions are equivalent:
\begin{enumerate}[(i)]
\item there exists a set~$M\dual$ of points of~$\surface$ and a cellular dissection~$\dissection\dual$ of~$(\surface, M\dual)$ such that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections,
\item each face of~$\dissection$ has at most one boundary arc.
\end{enumerate}
Moreover, the set~$M\dual$ and the cellular dissection~$\dissection\dual$ are both uniquely determined.
\end{proposition}

\begin{proof}
We mimick the proof of~\cite[Proposition 1.12]{OpperPlamondonSchroll}.
Assume first that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections. If a face~$F$ of~$\dissection$ had two boundary arcs~$a, b$, then it would contain two points of~$M\dual$, one on~$a$ and one on~$b$.
Conversely, if each face of~$\dissection$ has at most one boundary arc, we construct the set of points~$M\dual$ with one point on each boundary arc of~$\surface$ and one point in the interior of each interior face of~$\dissection$. Clearly, $M\dual$ has one point~$f\dual$ in each face~$f$ of~$\dissection$. We then construct half-arcs of~$\dissection\dual$ in each face~$f$ of~$\dissection$ by joining the dual point~$f\dual$ to the middle of each boundary arc of~$f$ that is not a boundary arc of~$\surface$. For each arc~$a$ of~$\dissection$, the half-arcs incident to the middle of~$a$ in the two faces of~$\dissection$ containing~$a$ form the dual arc~$a\dual$ of~$a$. All these dual arcs do not cross (as the half-arcs do not cross in each face of~$\dissection$) and form the dual cellular dissection~$\dissection\dual$ of~$\dissection$.
More details can be found in~\cite{OpperPlamondonSchroll}.
\end{proof}

@@

\subsection{Accordion complex, slalom complex, and non-crossing complex}

\begin{definition}
\label{def:accordion}
A \defn{$\dissection$-accordion} is an arc~$\gamma$ of~$(\surface, B)$ such that whenever~$\gamma$ crosses a face~$f$ of~$\dissection$, it crosses two edges of~$f$ that share a marked point of~$v \in M$ in such a way that~$v$ and~$f\dual$ lie on opposite sides of~$\gamma$.

@@

\begin{definition}
\label{def:slalom}
A \defn{$\dissection\dual$-slalom} is an arc~$\alpha$ of~$(\surface, B)$ such that, for each pair of faces~$(f,g) \in \faces(\dissection\dual)$ that~$\alpha$ enters consecutively by crossing~$\beta\in\dissection\dual$,~$\alpha$ cuts the union of the two cells $f$ and $g$ glued along~$\beta$ into two components, one containing the marked point $f\dual$ and the other containing $g\dual$.
\end{definition}

>>>>

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, xargs, multicol, overpic, pdflscape, multirow, hvfloat, array, ae, aecompl, pifont, mathtools, a4wide}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
%\usepackage[bottom]{footmisc}
\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles}
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}

@@

\begin{definition}
A \defn{marked surface}~$\bar\surface \eqdef (\surface, M)$ is an orientable surface~$\surface$ with boundaries, together with a set~$M$ of marked points which can be on the boundary of~$\surface$ or not.
For~$V \subset \surface$, a \defn{$V$-arc} on~$\bar\surface$ is a curve on~$\surface$ connecting two points of~$V$ and whose interior is disjoint from~$M$ and the boundary of~$\surface$.
As usual, arcs are considered up to homotopy relative to their endpoints in~$\surface \ssm M$, and curves homotopic to a boundary are not allowed.
Two arcs \defn{cross} when they intersect in~their~interior.
\end{definition}

\begin{definition}
A \defn{dissection} of~$\bar\surface$ is a collection~$\dissection$ of pairwise non-crossing arcs on~$\bar\surface$.
The \defn{faces} of a dissection~$\dissection$ are the connected components of the complement of the union of the arcs of~$\dissection$ in the surface~$\surface$.
We denote by~$\faces(\dissection)$ the set of faces of~$\dissection$.
The dissection~$\dissection$ is \defn{cellular} if all its faces are disks.
For~$V \subseteq M$, a \defn{$V$-dissection} is a dissection with only $V$-arcs.
\end{definition}

@@

\begin{definition}
Consider a marked surface~$\bar\surface = (\surface, V \sqcup V\dual)$, where~$V$ and~$V\dual$ are two disjoint sets of marked points so that the points of~$V$ and~$V\dual$ that are on the boundary of~$\surface$ alternate.
A cellular $V$-dissection~$\dissection$ of~$\bar\surface$ and a cellular $V\dual$-dissection~$\dissection\dual$ of~$\bar\surface$ are \defn{dual cellular dissections} if there are pairs of mutually inverse bijections~$V\dual \leftrightarrow \faces(\dissection)$ and~$V \leftrightarrow \faces(\dissection\dual)$, denoted~$\dual$ in both directions, such that~$\dissection$ has an arc joining its vertices~$u,v \in V$ and separating its faces~$f,g \in  \faces(\dissection)$ if and only if~$\dissection\dual$ has an arc joining its vertices~$f\dual, g\dual \in V\dual$ and separating its faces~$u\dual, v\dual \in \faces(\dissection\dual)$.
\end{definition}

Note that contrarily to the usual conventions, the dual vertex~$f\dual$ of a face~$f$ of~$\dissection$ is not always in the interior of the face~$f$.
More precisely, there are two situations:
\begin{itemize}
\item if a face~$f$ has no arc on the boundary of~$\surface$, its dual vertex~$f\dual$ lies in the interior~of~$f$,
\item if a face~$f$ has an arc on the boundary of~$\surface$, then it has exactly one such arc and its dual vertex~$f\dual$ lies on this arc.
\end{itemize}
In fact, the second point forces the following characterization of the cellular dissections that admit a dual cellular dissection.

\begin{proposition}
\label{prop:conditionsDualDissections}
For a cellular $V$-dissection~$\dissection$ of a marked surface~$\bar\surface = (\surface, V \sqcup V\dual)$, the following assertions are equivalent:
\begin{enumerate}[(i)]
\item there exists a cellular $V\dual$-dissection~$\dissection\dual$ of~$\bar\surface$ such that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections,
\item each face of~$\dissection$ contains exactly one point of~$V\dual$.
\end{enumerate}
Moreover, the cellular dissection~$\dissection\dual$ is uniquely determined.
\end{proposition}

\begin{proof}
We mimick the proof of~\cite[Prop.~1.12]{OpperPlamondonSchroll}.
%Assume first that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections. If a face~$f$ of~$\dissection$ had two boundary arcs~$a, b$, then it would contain two points of~$M\dual$, one on~$a$ and one on~$b$.
%Conversely, if each face of~$\dissection$ has at most one boundary arc, we construct the set of points~$M\dual$ with one point on each boundary arc of~$\surface$ and one point in the interior of each interior face of~$\dissection$. Clearly, $M\dual$ has one point~$f\dual$ in each face~$f$ of~$\dissection$. We then construct half-arcs of~$\dissection\dual$ in each face~$f$ of~$\dissection$ by joining the dual point~$f\dual$ to the middle of each boundary arc of~$f$ that is not a boundary arc of~$\surface$. For each arc~$a$ of~$\dissection$, the half-arcs incident to the middle of~$a$ in the two faces of~$\dissection$ containing~$a$ form the dual arc~$a\dual$ of~$a$. All these dual arcs do not cross (as the half-arcs do not cross in each face of~$\dissection$) and form the dual cellular dissection~$\dissection\dual$ of~$\dissection$.
%More details can be found in~\cite{OpperPlamondonSchroll}.
The direct implication is immediate by definition.
Assume conversely that each face~$f$ of~$\dissection$ contains exactly one point~$f\dual$ of~$V\dual$.
We construct half-arcs of~$\dissection\dual$ in each face~$f$ of~$\dissection$ by joining the dual point~$f\dual$ to the middle of each boundary arc of~$f$ that is not a boundary arc of~$\surface$.
For each arc~$a$ of~$\dissection$, the half-arcs incident to the middle of~$a$ in the two faces of~$\dissection$ containing~$a$ form the dual arc~$a\dual$ of~$a$.
All these dual arcs do not cross (as the half-arcs do not cross in each face of~$\dissection$) and form the dual cellular dissection~$\dissection\dual$~of~$\dissection$.
\end{proof}

@@

\subsection{Accordion complex, slalom complex, and non-crossing complex}
Let~$\surface$ be a surface with two disjoint sets~$V$ and~$V\dual$ of marked points, let~$B$ be the corresponding blossoming points, and consider the marked surface~$\bar\surface = (\surface, V \sqcup V\dual \sqcup B)$.
Consider two dual cellular dissections~$\dissection$ and~$\dissection\dual$ of~$\bar\surface = (\surface, V \sqcup V\dual)$.
%Let~$I$ denote the interior points and $B$ the blossom points.

\begin{definition}
\label{def:accordion}
A \defn{$\dissection$-accordion} is a $B$-arc~$\gamma$ of~$\bar\surface$ such that whenever~$\gamma$ crosses a face~$f$ of~$\dissection$, it crosses two edges of~$f$ that share a marked point of~$v \in M$ in such a way that~$v$ and~$f\dual$ lie on opposite sides of~$\gamma$.

@@

\begin{definition}
\label{def:slalom}
A \defn{$\dissection\dual$-slalom} is a $B$-arc~$\alpha$ of~$\bar\surface$ such that, for each pair of faces~$(f,g) \in \faces(\dissection\dual)$ that~$\alpha$ enters consecutively by crossing~$\beta\in\dissection\dual$,~$\alpha$ cuts the union of the two cells $f$ and $g$ glued along~$\beta$ into two components, one containing the marked point $f\dual$ and the other containing $g\dual$.
\end{definition}
